Learning by Teaching a Robot: the Case of Handwriting

### 2015-07-09 main.tex
A bit of cleaning

\subsection{Teaching beyond ICT}

This research departs from the traditional approaches to educative robotics
\cite{barker2007robotics,karim2015review}.

Also notable, the robot is not used in the usual context of robotics or computer
education, but instead in an activity -- handwriting -- which requires fine
physical skills. In such activities, the embodied nature of the robot is
appropriate as in interventions where motor mimicry is elicited
\cite{Berninger1997} the arm motion for instance is, \emph{by itself}, part of
the teaching. Furthermore, when facing a child with school difficulties, robots
can play the role of a na\"ive learner which neither adults nor peers -- because
of the social effects it would induce -- can convincingly play. Along these
lines, we hope to see more research on non-STEM educational applications of
robotics.

@@

Then, the actual implementation on the robot required the coordination of
several modules (from performing gestures and acquiring the user's input to
the high-level state machine), spread over several machines (the robot itself,
one laptop and up to four tactile tablets for certain studies we conducted). We
relied on ROS to ensure the synchronization and communication between these
modules.

Finally, the interaction itself, involving a robot, one or two children and, in
one study, an occupational therapist, also raised by itself several challenges.

We detail each of these points in the following sections.

@@

The main insight for both the generation and the learning of letters is to
reason in the eigenspace of the letters, instead of the cartesian space. The
eigen space of each letters is spanned by the first $n$ eigenvectors (in our
experiments, $3 < n < 6$) of the covariance matrix observed in a standard
dataset of adult letters (the UJI Pen Characters 2 dataset~\cite{Llorens2008}).
This procedure, based on a Principle Component Analysis (PCA), is explained in
details in a previously published article~\cite{hood2015when}.

@@

%
%\begin{figure}[ht]
%\centering
%
%\resizebox{0.9\linewidth}{!}{%
%
%\begin{tikzpicture}[
%    >=latex,
%    node distance=2cm,
%    every edge/.style={draw, very thick},
%    redarrow/.style={fill=Red, draw=Red},
%    greenarrow/.style={fill=GreenYellow, draw=GreenYellow},
%    yellowarrow/.style={fill=BurntOrange, draw=BurntOrange},
%    cmpt/.style={draw, align=center, rounded corners, inner sep=5pt, font=\sf, fill=black!20},
%    label/.style={midway, align=left, font=\scriptsize\sf, fill=white, above,opacity=0,text opacity=1}]
%
%    \node at (0,0)[cmpt, fill=Cyan!50, text width=2.7cm] (tactile) {tactile interaction \\ manager};
%    \node[cmpt, fill=Cyan!50, left=of tactile] (display) {display \\ manager};
%  
%    \node [cmpt, fill=Green!50, below=of tactile, text width=2.7cm] (shape) {shape position \\ manager};
%    \node [cmpt, fill=Green!50, left=of shape] (learner) {letter \\ learner};
%    \node [cmpt, fill=Green!50, left=of learner] (card) {card \\ detector};
%
%    \node [cmpt, fill=YellowOrange!50, below=1.5cm of shape] (writer) {nao writer};
%    \node [cmpt, fill=YellowOrange!50, left=of writer] (tablet) {tablet detector};
%    \node [cmpt, fill=YellowOrange!50, left=of tablet] (camera) {nao camera};
%    \node [cmpt, fill=YellowOrange!50, below=0.6cm of tablet] (pose) {robot state publisher};
%
%    %%% Relations between components
%    \path (tactile) edge [->] node[label] {clear \\ screen} (display);
%
%    \path ($(shape.north west)!0.33!(shape.north east)$) edge [<-] node[label,left] {clear \\ screen} ($(tactile.south west)!0.33!(tactile.south east)$);
%    \path ($(shape.north west)!0.66!(shape.north east)$) edge [<-, greenarrow] node[label,right] {feedback \\ + location} ($(tactile.south west)!0.66!(tactile.south east)$);
%
%    \path (shape) edge [->, redarrow] node[label, left, align=left] {letter \\to write \\ + time \\ + position} (display);
%    \path (shape) edge [->,redarrow] node[label, right] {letter to write \\ + time \\ + position in\\tablet frame} (writer);
%
%    \path ($(shape.north west)!0.66!(shape.south west)$) edge [->, greenarrow] node[label,below] {feedback \\ + letter} ($(learner.north east)!0.66!(learner.south east)$);
%    \path ($(shape.north west)!0.33!(shape.south west)$) edge [<-, redarrow] node[label] {letter to write \\ + time} ($(learner.north east)!0.33!(learner.south east)$);
%
%    \path (card) edge [->, yellowarrow] node[label] {word to learn} (learner);
%    \path (camera) edge [->, yellowarrow] node[label, right] {camera image} (card);
%    \path (camera) edge [->, redarrow] node[label] {camera image} (tablet);
%    \path (tablet) edge [->, redarrow] node[label] {tablet pose} (writer);
%    \path (pose) edge [->, redarrow] node[label, right] {head pose} (tablet);
%
%    \path (-8, 0) edge [->, redarrow] node[label] {Writing letters} (-7, 0);
%    \path (-8, -0.6) edge [->, greenarrow] node[label] {Responding to feedback} (-7, -0.6);
%    \path (-8, -1.2) edge [->, yellowarrow] node[label] {Detecting card input} (-7, -1.2);
%    
%\end{tikzpicture}
%}
%
%\caption{\small Overview of the system. Components in the top row run on the tablet, 
%those in the middle row on the central controller, and those in the bottom row 
%on the robot.}
%
%    \label{fig:archi}
%\end{figure}
%

@@

\end{itemize}
%
%\begin{figure}[thpb]
%     \begin{center}
%            \includegraphics[width=0.7\linewidth]{naoWriting2.png}
%    \end{center}
%    \caption{\small A demonstration of the robot simulating the writing of a word with
%    its finger. The motion of the robot is synchronised with the display of the
%    tablet, communicating over ROS.}%
%
%   \label{fig:naoWriting}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Interaction Design}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_distances}
    \caption{\small Two metrics to assess the handwriting progresses: Euclidian
    distance in the Eigenspace of the number dataset (top figure) or in
cartesian space (bottom figure). Green lines represent the robot performance,
blue lines Henry's performance. The round IDs correspond to the demonstrations
pictured on Figure~\ref{henry_demos}.}
    \label{henry_distances}
\end{figure}

@@

No surprinsingly, the role of the robot \textit{vis-a-vis} the teacher has also
been questionned (including an angry email from one teacher: ``You want to
replace us!''): as we see it, the role of the robot within the classroom (or at
the therapist's surgery) does not infringe upon the role of the adult (teacher
or therapist).  The core of the \emph{learning by teaching} paradigm relies on
the child becoming the teacher of an underperforming pupil (the robot): from
that perspective, the robot does not replace the teacher, on the contrary. It
plays a different role in the classroom, which happen to be novel as well: the
robot is the \emph{least} performing student, and however a very patient, always
eager to improve, one.

@@

We initially envisionned the CoWriter system to be run in the back of a
classroom with one child: this would have allowed an individual, face-to-face
remediation approach, not otherwise tractable for a teacher with 20 pupils.
This is unlikely to happen. In our experience, the teacher keeps such an
important role that the interaction and the learning would hardly occur if the
child is left alone (or even semi-alone). At the end, we can reassure teachers:
robots are not going to replace them any time soon.

>>>>

@@

Then, the actual implementation on the robot required the coordination of
several modules (from performing gestures and acquiring the user's input to
the high-level state machine), spread over several devices (the robot itself,
one laptop and up to four tactile tablets for certain studies we conducted). We
relied on ROS to ensure the synchronization and communication between these
modules.

We detail each of these in the following sections.

@@

The main insight for both the generation and the learning of letters is to
reason in the eigenspace of the letters, instead of the cartesian space. The
eigenspace of each letters is spanned by the first $n$ eigenvectors (in our
experiments, $3 < n < 6$) of the covariance matrix observed in a standard
dataset of adult letters (the UJI Pen Characters 2 dataset~\cite{Llorens2008}).
This procedure, based on a Principle Component Analysis (PCA), is explained in
details in a previously published article~\cite{hood2015when}.

@@

@@

\end{itemize}

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_distances}
    \caption{\small Two metrics to assess the handwriting progresses: Euclidian
    distance in the eigenspace of the number dataset (top figure) or in
cartesian space (bottom figure). Green lines represent the robot performance,
blue lines Henry's performance. The round IDs correspond to the demonstrations
pictured on Figure~\ref{henry_demos}.}
    \label{henry_distances}
\end{figure}

@@

No surprisingly, the role of the robot \textit{vis-a-vis} the teacher has also
been questioned (including an angry email from one teacher: ``You want to
replace us!''): as we see it, the role of the robot within the classroom (or at
the therapist's surgery) does not infringe upon the role of the adult (teacher
or therapist).  The core of the \emph{learning by teaching} paradigm relies on
the child becoming the teacher of an underperforming pupil (the robot): from
that perspective, the robot does not replace the teacher, on the contrary. It
plays a different role in the classroom, which happen to be novel as well: the
robot is the \emph{least} performing student, and however a very patient, always
eager to improve, one.

@@

We initially envisioned the CoWriter system to be run in the back of a
classroom with one child: this would have allowed an individual, face-to-face
remediation approach, not otherwise tractable for a teacher with 20 pupils.
This is unlikely to happen. In our experience, the teacher keeps such an
important role that the interaction and the learning would hardly occur if the
child is left alone (or even semi-alone). At the end, we can reassure teachers:
robots are not going to replace them any time soon.
